---
title: "Week 4 DS Recitation 2: Tidy Data"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "SIBDS 2023@ Columbia" 
output: 
  html_document:
    toc: true
    code_folding: "show"
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
set.seed(20230614)
library(tidyverse)

# set knitr defaults
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 6,
  fig.asp   = .6, # if specified, fig.height = fig.width * fig.asp
  out.width = "90%",
  cache = FALSE
)

# set theme defaults
theme_set(
  theme_bw() +
    theme(
      legend.position = "bottom",
      plot.title    = element_text(hjust = 0.5),
      plot.subtitle = element_text(hjust = 0.5),
      plot.caption  = element_text(hjust = 0.0)
    )
)

# set color scale defaults
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill   = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete   = scale_fill_viridis_d
```

## Getting Started

### Tasks:

  1. Put the `week_4_DS_recitation_2_tidy_data.Rmd` file into the same folder of the R project you created. 

## Pivot wider

### Tasks:

  1. The following `warpbreaks` dataset contains the warp break experiment results with nine replicates for every combination of `wool` (`A` and `B`) and `tension` (`L`, `M`, `H`):
  
```{r}
warpbreaks
```
Try to run the following code, what happens if we attempt to pivot the levels of `wool` into the columns?

```{r}
warpbreaks %>% 
  pivot_wider(
    names_from = wool, 
    values_from = breaks
  )
```
Try to change the default setting of `values_fn` to get a wider dataframe summarizing the `mean` of those 9 experiment results (mean `breaks` for each combination of `wool` and `tension`).

```{r}
# Try to modify this code
warpbreaks %>% 
  pivot_wider(
    names_from = wool, 
    values_from = breaks,
    values_fn = mean
  )
```

  2. The `us_rent_income` dataset contains information about median income and rent for each state in the US for 2017. Here both `estimate` and `moe` (1.645 $\times$ SE) are values columns. Try to provide a summary of the data for each state in a single row, outlining both income and rent estimates as well as their variances.
  
```{r}
us_rent_income
# Your answer starts here
us_rent_income %>% 
  pivot_wider(
    names_from = variable, 
    values_from = c(estimate, moe)
  )
```
It’s relatively rare to need `pivot_wider()` to make tidy data, but it’s often useful for creating summary tables for presentation, or data in a format needed by other tool
 

## Longer then wider

### Tasks:

1. The `world_bank_pop` dataset contains data from the World Bank about population per country from 2000 to 2018.

```{r}
world_bank_pop
```
Try to tidy the data, you can follow these steps:

  - Firstly, `year` is spread across multiple columns, we can use `pivot_longer` to put them into a single column
  
  - Next, focus on the `indicator`. Here `SP.POP.GROW` is population growth, `SP.POP.TOTL` is total population, and `SP.URB.*` are the same but only for urban areas. Let’s split this up into two variables: `area` (total or urban) and the actual `variable` (population or growth), you may need this : `separate(indicator, c("SP", "area", "variable"))`

  - Finally, we can complete the tidying by pivoting (using `pivot_wider`) `variable` and `value` to make `TOTL` and `GROW` columns


```{r}
# Your answer starts here
world_bank_pop %>% 
  # aggregate years to one column
  pivot_longer(
    cols = `2000`:`2017`, 
    names_to = "year", 
    values_to = "value"
  ) %>% 
  # separate the indicator to 3 variables and drop the first one
  separate(indicator, c(NA, "area", "variable")) %>% 
  # summarize the 
  pivot_wider(
    names_from = variable, 
    values_from = value
  )
```


## Joining datasets

### Tasks:

  1. Install and load the `nycflights13` dataset use the following code. 

```{r}
if (!requireNamespace("nycflights13", quietly = TRUE)) {
  install.packages("nycflights13")
}

library(nycflights13)
```

  2. Join the `weather` dataset to the `flights2` dataset (created by the following code) using `left_join`. Which variables are used in this joining process?
  
```{r}
flights2 <- flights %>% select(year:day, hour, origin, dest, tailnum, carrier)
# Your answer starts here
flights2 %>% left_join(weather)
```
  
  3. Join the `airports` dataset to the `flights2` dataset using `left_join`. The variable we want to use in the `airports` dataset is `faa`. Noticing that each flight has an origin (`origin`) and destination (`dest`) airport, so we need to specify which one we want to join to.
  
```{r}
# Your answer starts here
# joining based on destination
flights2 %>% left_join(airports, c("dest" = "faa"))

# joining based on origin
flights2 %>% left_join(airports, c("origin" = "faa"))
```
