---
title: "Week 5 DS Recitation 1: Data Visualization"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "SIBDS 2023@ Columbia" 
output: 
  html_document:
    toc: true
    code_folding: "show"
---

```{r setup, include=FALSE}
set.seed(20230614)
library(tidyverse)

# set knitr defaults
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 6,
  fig.asp   = .6, # if specified, fig.height = fig.width * fig.asp
  out.width = "90%",
  cache = FALSE
)

# set theme defaults
theme_set(
  theme_bw() +
    theme(
      legend.position = "bottom",
      plot.title    = element_text(hjust = 0.5),
      plot.subtitle = element_text(hjust = 0.5),
      plot.caption  = element_text(hjust = 0.0)
    )
)

# set color scale defaults
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill   = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete   = scale_fill_viridis_d
```

## Things We Left

Here is a useful link of [ggplot cheatsheets](https://github.com/rstudio/cheatsheets/blob/main/data-visualization.pdf).

```{r}
# Using the weather_df as an example
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (ºC)",
    y = "Maxiumum daily temperature (ºC)",
    caption = "Data from the rnoaa package", 
    color = "Park") + 
  scale_x_continuous(
    breaks = seq(-15, 20, 5), 
    labels = paste0(seq(-15, 20, 5), "ºC", sep = " ")) +
  scale_fill_discrete(name = "Location", labels = c("Central Park, NY", "Waikiki, HA", "Waterhole, WA"), aesthetics = "color")


```

- You can use `seq` to generate a sequence of numbers as ticks

- Use `scale_fill_discrete` to manually change the text in the legend. Using either `labs` or `scale_fill_discrete` (with the `aesthestics` argument) to change the title of the legend.


## Getting Started

  1. Create a new R project and named it `Week 5 DS Recitation`

  2. Put the `week_5_DS_recitation_1_data_viz.Rmd` file into the same folder of the R project you just created. 


## Visualization

We are going to use NY NOAA data. Please load the data from the `p8105.datasets` package using:

```{r}
if (!requireNamespace("p8105.datasets", quietly = TRUE)) {
  devtools::install_github("p8105/p8105.datasets")
}

library(p8105.datasets)
data("ny_noaa")
```

Here, I did some basic cleaning for you:

```{r}
noaa_df = ny_noaa %>% 
  separate(date, into = c("year","month","day"),sep = "-") %>% 
  mutate(
    year = as.factor(year),
    month = as.integer(month),
    month = (month.name[month]),
    day = as.factor(day),
    prcp = prcp / 10,
    tmax = as.numeric(tmax),
    tmin = as.numeric(tmin),
    tmax = tmax / 10,
    tmin = tmin / 10
  ) 

# Here is a table shows the most common observations of snowfall
noaa_df %>% 
  count(snow, name = "n_of_observations") %>% 
  drop_na(snow) %>% 
  arrange(desc(n_of_observations)) %>% 
  head(n = 5)
```
### Tasks:

Please complete the following tasks in groups of two or three:

  1. Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?
  
```{r}
# You may start with this processed data
noaa_df %>%
  #focus on Jan and Jul
  filter(month %in% c("January", "July")) %>% 
  group_by(year, month, id) %>% 
  summarise(mean_of_tmax = mean(tmax,na.rm = TRUE))
```
  

  2. Make a two-panel plot showing (i) `tmax` vs `tmin` for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.
  
Discuss with your group members to come up with the best way to visualize data.

