---
title: "Week 5 DS Recitation: Exploratory Data Analysis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "SIBDS 2024 @ Columbia" 
output: 
  html_document:
    toc: true
    code_folding: "show"
---

```{r setup, include=FALSE}
set.seed(2024)
library(tidyverse)

# set knitr defaults
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 6,
  fig.asp   = .6, # if specified, fig.height = fig.width * fig.asp
  out.width = "90%",
  cache = FALSE
)

# set theme defaults
theme_set(
  theme_bw() +
    theme(
      legend.position = "bottom",
      plot.title    = element_text(hjust = 0.5),
      plot.subtitle = element_text(hjust = 0.5),
      plot.caption  = element_text(hjust = 0.0)
    )
)

# set color scale defaults
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill   = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete   = scale_fill_viridis_d
```


## Getting Started
  
  1. Create a new R project and named it `week5_DS_recitation`
  
  2. Put the `week5_DS_recitation_data_viz_2.Rmd` file into the same folder of the R project you created. 


## EDA with `nycflights13`

We are going to use the `nycflights13` package again. Please load the `flights` data from the `nycflights13` package using:

```{r}
if (!requireNamespace("nycflights13", quietly = TRUE)) {
  install.packages("nycflights13")
}

library(nycflights13)
data("flights")
```

### Tasks:

1. Which plane (`tailnum`) has the worst on-time record (remove planes that flew `<` 20 flights)? To find out, you may create an indicator to determine the on-time record with the code `mutate(on_time = !is.na(arr_time) & (arr_delay <= 0))`, or you can consider the average number of minutes delayed with `mean(arr_delay)`.

```{r}
# Your answer starts here
flights %>%
  filter(!is.na(tailnum))
# Your answer starts here
flights %>% 
  filter(!is.na(tailnum), is.na(arr_time) | !is.na(arr_delay))
```



2. What time of day (`hour`) should you fly if you want to avoid delays as much as possible?

```{r}
# Your answer starts here
```


3. Delays are typically temporally correlated: even once the problem that caused the initial delay has been resolved, later flights are delayed to allow earlier flights to leave. Using `lag()`, explore how the delay of a flight is related to the delay of the immediately preceding flight, please use a plot to display the relationship.
.

```{r}
# Your answer starts here
```



