---
title: "Week 3 DS Recitation"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "SIBDS 2023@ Columbia"
output: pdf_document
---

```{r setup, include=FALSE}
set.seed(20230614)
library(tidyverse)

# set knitr defaults
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 6,
  fig.asp   = .6,
  out.width = "90%",
  cache = TRUE
)

# set theme defaults
theme_set(
  theme_bw() +
    theme(
      legend.position = "bottom",
      plot.title    = element_text(hjust = 0.5),
      plot.subtitle = element_text(hjust = 0.5),
      plot.caption  = element_text(hjust = 0.0)
    )
)

# set color scale defaults
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill   = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete   = scale_fill_viridis_d
```

## Getting Started

### Tasks:

1. Create a new R project and named it `Week 3 DS Recitation`

2. Put the `week_3_DS_recitation.Rmd` file into the same folder of the R project you just created. 

3. Modify the YAML header as follows:

```
---
title: "Week 3 DS Recitation"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "SIBDS 2023@ Columbia" 
output: 
  html_document:
    toc: true
    code_folding: "show"
---
```

   Now, knit this `.Rmd` file to make sure it runs as expected.

4. Finally, create a folder within the project folder named `data`. Put the data files we downloaded in Tuesday's class in to this folder. 

## Writing with Data

### Tasks:

  1. Write a named code chunk names `sample_df` that creates a dataframe comprised of
  
      * `random_vec` : A `numeric` variable containing a random sample of size `500` from a normal variable with mean `1` and standard deviation `3`.
    
      * `logical_vec` : A `logical` vector indicating whether each sampled value is greater than zero, can you do it in the simplest way?
    
      * `abs_vec`: A `numeric` vector containing the absolute value (using `abs()`) of each element. 
  
```{r}
# Your answer starts here:  
```
  

  2. Produce a histogram of the absolute value variable just created.
  
```{r}
# Your answer starts here:  
```
  


  4. Fill in the blank using the in-line code giving the mean, median, and standard deviation of `random_vec` (*Hint*: you may need to use the `pull()` function in `tidyverse`):
  
      * Mean: `REPLACE THIS BY YOUR CODE`
  
      * Median: `REPLACE THIS BY YOUR CODE`   
  
      * SD: `REPLACE THIS BY YOUR CODE`

## Data Import

### Tasks:

  1. Load the `FAS_pups.csv` dataset. Use both absolute and relative paths
  
```{r}
# Your answer starts here:  
```
  

  
















