---
title: 'Week 3 DS Recitation Part 2'
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "SIBDS 2024 @ Columbia"
output: pdf_document
---

## Data Import

### Tasks:

  1. Load the `FAS_pups.csv` dataset. Use both absolute and relative paths.
  
```{r data_import}
# Your answer starts here:  

```

  2. Import the `FAS_pups.csv` again with column specifications by setting the `col_type` arguments. What happens if your specifications for column parsing aren’t reasonable (e.g. character instead of double, or vice versa)? 
  
```{r}
# Let's see what happens together: 

```


## Data Manipulation

### Tasks:

  1. In the pups data, do the following:
      a. select the columns containing litter number, sex, and PD ears
      b. Filter to include only pups with `sex == 1`
      c. Filter to include only pups with `pd_walk` less than 11 and `sex == 2`
      d. Create a variable that subtracts 7 from `pd_pivot`
      e. Create a variable that is the sum of all the `pd` variables
      f. Sort the variable `pd_eyes` in ascending order and `pd_walk` in descending order
      
```{r}
# Your answer starts here:  

```
      
  2. Write a chain of commands that:
      - Loads the `pups` data from `FAS_pups.csv`
      - Cleans the variable names
      - Filters the data to include only `pups` with `sex == 1`
      - Removes the `pd_ears` variable
      - Creates a variable that indicates whether `pd_pivot` is 7 or more days
```{r}
# Your answer starts here:  

```
















  