---
title: "Week 6 DS Recitation: Regression"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "SIBDS 2023@ Columbia" 
output: 
  html_document:
    toc: true
    code_folding: "show"
---

```{r setup, include=FALSE}
set.seed(20230614)
library(tidyverse)

# set knitr defaults
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 6,
  fig.asp   = .6, # if specified, fig.height = fig.width * fig.asp
  out.width = "90%",
  cache = FALSE
)

# set theme defaults
theme_set(
  theme_bw() +
    theme(
      legend.position = "bottom",
      plot.title    = element_text(hjust = 0.5),
      plot.subtitle = element_text(hjust = 0.5),
      plot.caption  = element_text(hjust = 0.0)
    )
)

# set color scale defaults
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill   = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete   = scale_fill_viridis_d
```


## Getting Started

  1. Create a new R project and named it `Week 6 DS Recitation`

  2. Put the `week_6_DS_recitation_regression.Rmd` file and the `birthweight.csv` into the same folder of the R project you just created.


## Linear Regression

In this problem, you will analyze data gathered to understand the effects of several variables on a child’s birth weight. This dataset (`birthweight.csv`), consists of roughly 4000 children and includes the following variables:

- babysex: baby’s sex (male = 1, female = 2)
- bhead: baby’s head circumference at birth (centimeters)
- blength: baby’s length at birth (centimeteres)
- bwt: baby’s birth weight (grams)
- delwt: mother’s weight at delivery (pounds)
- fincome: family monthly income (in hundreds, rounded)
- frace: father’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other, 9 = Unknown)
- gaweeks: gestational age in weeks
- malform: presence of malformations that could affect weight (0 = absent, 1 = present)
- menarche: mother’s age at menarche (years)
- mheigth: mother’s height (inches)
- momage: mother’s age at delivery (years)
- mrace: mother’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other)
- parity: number of live births prior to this pregnancy
- pnumlbw: previous number of low birth weight babies
- pnumgsa: number of prior small for gestational age babies
- ppbmi: mother’s pre-pregnancy BMI
- ppwt: mother’s pre-pregnancy weight (pounds)
- smoken: average number of cigarettes smoked per day during pregnancy
 -wtgain: mother’s weight gain during pregnancy (pounds)
 
Load and clean the data using the code below for regression analysis.

```{r}
library(tidyverse)
baby_df = 
  read_csv("./birthweight.csv") %>% 
  mutate(
    babysex = case_when(babysex == 1~"male", babysex == 2~"female"),
    malform = case_when(malform == 0~"absent", malform == 1 ~ "present")
    ) %>% 
  mutate(
    across(
      .cols = c("frace", "mrace"),
      ~ case_when(
        .x == 1 ~ "White",
        .x == 2 ~ "Black",
        .x == 3 ~ "Asian",
        .x == 4 ~ "Puero Rican",
        .x == 8 ~ "Other",
        .x == 9 ~ "Unknown"
        )
      ),
    across(where(is.character), as.factor)
    ) 
```


1. Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birth weight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use `add_predictions` and `add_residuals` in making this plot.

```{r}
# Your answer starts here
proposal_model = 
  lm(bwt ~ babysex + blength + gaweeks + momage + parity + ppwt + wtgain + smoken, data = baby_df)

proposal_model %>% broom::tidy() %>% knitr::kable(digits = 3)

proposal_plt =
  baby_df %>% 
  add_predictions(proposal_model) %>%
  add_residuals(proposal_model) 

library(patchwork)
pred_plt1 = 
  proposal_plt %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .1)

pred_plt2 = 
  proposal_plt %>% 
  ggplot(aes(x = bwt, y = pred)) +
  geom_point(alpha = .1)

pred_plt3 = 
  proposal_plt %>% 
  ggplot(aes(x = resid))+
  geom_density()

pred_plt1/pred_plt2/pred_plt3
```

2. (Chanllenging) Compare your model to two others:

- One using length at birth and gestational age as predictors (main effects only)
- One using head circumference, length, sex, and all interactions (including the three-way interaction) between these   

```{r}
# Your answer starts here
# Take a look at the regressions first
my_model = lm(bwt ~ babysex + blength + gaweeks + momage + parity + ppwt + wtgain + smoken, data = baby_df)
simple_model = lm(bwt ~ bhead + gaweeks, data = baby_df)
complex_model = lm(bwt ~ bhead*blength*babysex, data = baby_df)

baby_cv = 
  crossv_mc(baby_df, 100) %>% 
  mutate(
    my_model = map(train,  ~lm(bwt ~ babysex + blength + gaweeks + momage + parity + ppwt + wtgain + smoken, data = .x)),
    simple_model = map(train, ~lm(bwt ~ bhead + gaweeks, data = .x)),
    complex_model = map(train, ~lm(bwt ~ bhead*blength*babysex, data = .x)),
    rmse_my = map2_dbl(my_model, test, ~rmse(model = .x, data = .y)),
    rmse_simple = map2_dbl(simple_model, test, ~rmse(model = .x, data = .y)),
    rmse_complex = map2_dbl(complex_model, test, ~rmse(model = .x, data = .y))
  )

baby_cv_tidy = 
  baby_cv %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  mutate(model = fct_inorder(model))
  

baby_cv_tidy %>% 
  ggplot(aes(x = model, y = rmse, group = model, color = model)) +
  geom_violin(alpha = .3) +
  theme(legend.position = "none") +
  labs(
    title = "RMSE for Three Models",
    x = "Models",
    y = "RMSE"
  )
```

