---
title: "Week 5 DS Recitation 2: Exploratory Data Analysis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "SIBDS 2023@ Columbia" 
output: 
  html_document:
    toc: true
    code_folding: "show"
---

```{r setup, include=FALSE}
set.seed(2024)
library(tidyverse)

# set knitr defaults
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.asp   = .6, # if specified, fig.height = fig.width * fig.asp
  out.width = "90%",
  cache = FALSE
)

# set theme defaults
theme_set(
  theme_bw() +
    theme(
      legend.position = "bottom",
      plot.title    = element_text(hjust = 0.5),
      plot.subtitle = element_text(hjust = 0.5),
      plot.caption  = element_text(hjust = 0.0)
    )
)

# set color scale defaults
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill   = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete   = scale_fill_viridis_d
```


## Getting Started

  1. Put the `week_5_DS_recitation_2_EDA.Rmd` file into the same folder of the R project you created on Wednesday. 


## EDA with `nycflights13`

We are going to use the `nycflights13` package again. Please load the `flights` data from the `nycflights13` package using:

```{r}
if (!requireNamespace("nycflights13", quietly = TRUE)) {
  install.packages("nycflights13")
}

library(nycflights13)
data("flights")
```

### Tasks:

1. Which plane (`tailnum`) has the worst on-time record (remove planes that flew `<` 20 flights)? To find out, you may create an indicator to determine the on-time record with the code `mutate(on_time = !is.na(arr_time) & (arr_delay <= 0))`, or you can consider the average number of minutes delayed with `mean(arr_delay)`.

```{r}
# Your answer starts here
flight1 = 
  flights %>%
  filter(!is.na(tailnum), !is.na(arr_time), !is.na(arr_delay)) %>%
  mutate(on_time = !is.na(arr_time) & (arr_delay <= 0)) %>%
  group_by(tailnum) %>%
  summarise(on_time = mean(on_time), n = n()) %>%
  filter(n >= 20) %>%
  filter(min_rank(on_time) == 1)
# N988AT
```


```{r}
# using mean arr_delay
flight2 = 
  flights %>%
  filter(!is.na(tailnum), !is.na(arr_time), !is.na(arr_delay)) %>% 
  group_by(tailnum) %>%
  summarise(arr_delay = mean(arr_delay), n = n()) %>%
  filter(n >= 20) %>%
  filter(min_rank(desc(arr_delay)) == 1)
# N203FRS
```


2. What time of day (`hour`) should you fly if you want to avoid delays as much as possible?

```{r}
# Your answer starts here
flights %>%
  group_by(hour) %>%
  summarise(arr_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(arr_delay)
# 7AM
```


3. Delays are typically temporally correlated: even once the problem that caused the initial delay has been resolved, later flights are delayed to allow earlier flights to leave. Using `lag()`, explore how the delay of a flight is related to the delay of the immediately preceding flight, please use a plot to display the relationship.

```{r}
# Your answer starts here
lagged_delays <- flights %>%
  arrange(origin, month, day, dep_time) %>%
  group_by(origin) %>%
  mutate(dep_delay_lag = lag(dep_delay)) %>%
  filter(!is.na(dep_delay), !is.na(dep_delay_lag))

lagged_delays %>%
  group_by(dep_delay_lag, origin) %>%
  summarise(dep_delay_mean = mean(dep_delay)) %>%
  ggplot(aes(y = dep_delay_mean, x = dep_delay_lag, color = origin)) +
  geom_point(alpha = 0.5) +
  scale_x_continuous(breaks = seq(0, 1500, by = 120)) +
  labs(y = "Mean Departure Delay", x = "Previous Departure Delay") +
  facet_wrap(.~origin)
```



