---
title: "Week 4 DS Recitation: Tidy Data"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "SIBDS 2024 @ Columbia" 
output: 
  html_document:
    toc: true
    code_folding: "show"
---

```{r setup, include=FALSE}
set.seed(20230614)
library(tidyverse)

# set knitr defaults
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 6,
  fig.asp   = .6, # if specified, fig.height = fig.width * fig.asp
  out.width = "90%",
  cache = FALSE
)

# set theme defaults
theme_set(
  theme_bw() +
    theme(
      legend.position = "bottom",
      plot.title    = element_text(hjust = 0.5),
      plot.subtitle = element_text(hjust = 0.5),
      plot.caption  = element_text(hjust = 0.0)
    )
)

# set color scale defaults
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill   = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete   = scale_fill_viridis_d
```

## Getting Started

### Tasks:

1. Create a new R project and named it `week4_DS_recitation`

2. Put the `week4_DS_recitation_tidy_data.Rmd` file into the same folder of the R project you just created. 

## Pivot longer

### Tasks:
  
1.The `billboard` dataset in `tidyr` package records the billboard rank of songs in the year 2000.
```{r}
billboard %>% head(10)
``` 
  
Reshape the dataset by changing the `wk1` to `wk76` to a variable called `week`, and the values to a variable called `rank`. Also try to use `values_drop_na` to drop rows that correspond to missing values (Not every song stays in the charts for all 76 weeks).   

```{r}
# Your answer starts here

```
  
  
2. Redo the above reshaping process, convert the `week` variable to an integer this time.
  
```{r}
# Your answer starts here

```


3. The following `household` dataset contains information (or values) for each child in different families: their `name` and their `dob` (date of birth).
```{r}
head(household)
```

As you can see, we have multiple records for some families since there is more than one child in that family. In this case, how can we tidy the data so that each row in the new dataset represents the record of a child in his/her family with the child's name, date of birth, and whether or not he/she is the first child? The following code give you the desired tidy dataset in a painful way.

```{r eval = F}


```


## Pivot wider

### Tasks:

1. The following `warpbreaks` dataset contains the warp break experiment results with nine replicates for every combination of `wool` (`A` and `B`) and `tension` (`L`, `M`, `H`):
  
```{r}
warpbreaks
```

Try to run the following code, what happens if we attempt to pivot the levels of `wool` into the columns?

```{r}
warpbreaks %>% 
  pivot_wider(
    names_from = wool, 
    values_from = breaks
  )
```

Try to change the default setting of `values_fn` to get a wider dataframe summarizing the `mean` of those 9 experiment results (mean `breaks` for each combination of `wool` and `tension`).

```{r}
# Try to modify this code
warpbreaks %>% 
  pivot_wider(
    names_from = wool, 
    values_from = breaks,
    values_fn = mean
  )
```

2. The `us_rent_income` dataset contains information about median income and rent for each state in the US for 2017. Here both `estimate` and `moe` (1.645 $\times$ SE) are values columns. Try to provide a summary of the data for each state in a single row, outlining both income and rent estimates as well as their variances.
  
```{r}
us_rent_income 
# Your answer starts here

```


 
## Longer then wider

### Tasks:

1. The `world_bank_pop` dataset contains data from the World Bank about population per country from 2000 to 2018.

```{r}
world_bank_pop
```
Try to tidy the data, you can follow these steps:

  - Firstly, `year` is spread across multiple columns, we can use `pivot_longer` to put them into a single column
  
  - Next, focus on the `indicator`. Here `SP.POP.GROW` is population growth, `SP.POP.TOTL` is total population, and `SP.URB.*` are the same but only for urban areas. Let’s split this up into two variables: `area` (total or urban) and the actual `variable` (population or growth), you may need this : `separate(indicator, c("SP", "area", "variable"))`

  - Finally, we can complete the tidying by pivoting (using `pivot_wider`) `variable` and `value` to make `TOTL` and `GROW` columns


```{r}
# Your answer starts here

```


## Joining datasets

### Tasks:

1. Install and load the `nycflights13` dataset use the following code. 

```{r}
if (!requireNamespace("nycflights13", quietly = TRUE)) {
  install.packages("nycflights13")
}

library(nycflights13)
```

2. Join the `weather` dataset to the `flights2` dataset (created by the following code) using `left_join`. Which variables are used in this joining process?
  
```{r}
flights2 <- flights %>% select(year:day, hour, origin, dest, tailnum, carrier)
# Your answer starts here

```
  
3. Join the `airports` dataset to the `flights2` dataset using `left_join`. The variable we want to use in the `airports` dataset is `faa`. Noticing that each flight has an origin (`origin`) and destination (`dest`) airport, so we need to specify which one we want to join to.
  
```{r}
# Your answer starts here
data(airports)

```


  