---
title: "Week 3 DS Recitation"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "SIBDS 2023@ Columbia" 
output: 
  html_document:
    toc: true
    code_folding: "show"
---

```{r setup, include=FALSE}
set.seed(20230614)
library(tidyverse)

# set knitr defaults
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 6,
  fig.asp   = .6, # if specified, fig.height = fig.width * fig.asp
  out.width = "90%",
  cache = FALSE
)

# set theme defaults
theme_set(
  theme_bw() +
    theme(
      legend.position = "bottom",
      plot.title    = element_text(hjust = 0.5),
      plot.subtitle = element_text(hjust = 0.5),
      plot.caption  = element_text(hjust = 0.0)
    )
)

# set color scale defaults
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill   = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete   = scale_fill_viridis_d
```

## Getting Started

### Tasks:

1. Create a new R project and named it `Week 3 DS Recitation`

2. Put the `week_3_DS_recitation.Rmd` file into the same folder of the R project you just created. 

3. Modify the YAML header as follows:

```
---
title: "Week 3 DS Recitation"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "SIBDS 2023@ Columbia" 
output: 
  html_document:
    toc: true
    code_folding: "show"
---
```

   Now, knit this `.Rmd` file to make sure it runs as expected.

4. Finally, create a folder within the project folder named `data`. Put the data files we downloaded in Tuesday's class in to this folder. 

## Writing with Data

### Tasks:

  1. Write a named code chunk names `sample_df` that creates a dataframe comprised of
  
      * `random_vec` : A `numeric` variable containing a random sample of size `500` from a normal variable with mean `1` and standard deviation `3`.
    
      * `logical_vec` : A `logical` vector indicating whether each sampled value is greater than zero, can you do it in the simplest way?
    
      * `abs_vec`: A `numeric` vector containing the absolute value (using `abs()`) of each element. 
  
```{r sample_df}
# Your answer starts here: 
df1 <- tibble(
random_vec = rnorm(500, mean = 1, sd = 3),
logical_vec = random_vec > 0,
abs_vec = abs(random_vec)
)
```
  

  2. Produce a histogram of the absolute value variable just created.
  
```{r sample_hist}
# Your answer starts here:  
ggplot(aes(x = abs_vec), data = df1) + geom_histogram()
#df1 %>% ggplot(aes(x = abs_vec)) + geom_histogram()
```
  


  4. Fill in the blank using the in-line code giving the mean, median, and standard deviation of `random_vec` (*Hint*: you may need to use the `pull()` function and  in `tidyverse`), round them to two decimal places.
  
      * Mean: `r round(mean(pull(df1, random_vec)),2)`
  
      * Median: `r round(median(pull(df1, random_vec)),2)`   
  
      * SD: `r round(sd(pull(df1, random_vec)),2)`

## Data Import

### Tasks:

  1. Load the `FAS_pups.csv` dataset. Use both absolute and relative paths
  
```{r data_import}
# Your answer starts here:  
working_dir = getwd()
data_dir = paste0(working_dir,"/data/FAS_pups.csv")
# absolute path
pups = read_csv(data_dir)
pups = janitor::clean_names(pups)
# relative path
pups = read_csv("./data/FAS_pups.csv")
pups = janitor::clean_names(pups)

# Take a quick look of your dataset
#summary(pups)

#gtsummary::tbl_summary(pups)

#summarytools::dfSummary(pups)
```
  
2. Import the `FAS_pups.csv` again with column specifications by setting the `col_type` arguments. What happens if your specifications for column parsing aren’t reasonable (e.g. character instead of double, or vice versa)? 
  
```{r}
# Let's see what happens together: 
pups_bad = read_csv("./data/FAS_pups.csv", col_types = "cccccc")
pups_bad = janitor::clean_names(pups_bad)
skimr::skim(pups_bad) # skim is smart

pups_really_bad = read_csv("./data/FAS_pups.csv", col_types = "iiiiii")
pups_really_bad = janitor::clean_names(pups_really_bad)
skimr::skim(pups_really_bad)
```


## Data Manipulation

### Tasks:

  1. In the pups data, select the columns containing litter number, sex, and PD ears. Then, do the following:
      a. Filter to include only pups with `sex == 1`
      b. Filter to include only pups with `pd_walk` less than 11 and `sex == 2`
      c. Create a variable that subtracts 7 from `pd_pivot`
      d. Create a variable that is the sum of all the `pd` variables
```{r}
# Your answer starts here:  
# filter a
pups %>% filter(sex == 1)

# filter b
pups %>% filter(pd_walk < 11 & sex == 2)

# filter c
# base R style
pups$pd_pivot_7 = pups$pd_pivot - 7
pups["pd_pivot_7_2"] = pups["pd_pivot"] - 7
#pups[,9] = pups[,5] - 7 Better not

# using pipe operator
pups_c <- pups %>%
  mutate(pd_pivot_7 = pd_pivot - 7) %>%
  relocate(pd_pivot_7)
pups_c

# filter d  
pups_d <-
  pups %>%
  mutate(
    pd_sum = pd_ears + pd_eyes + pd_pivot + pd_walk
) 
pups_d

pups_d_na_rm <- 
  pups %>%  
  mutate(
    pd_sum = rowSums(
      select(., starts_with("pd_")), 
      na.rm = TRUE)
    )
pups_d_na_rm

#rowSums(
#  select(pups, starts_with("pd_"), -ends_with("7")), 
#  na.rm = TRUE
#)
```
      
      
  2. Write a chain of commands that:
      - Loads the `pups` data from `FAS_pups.csv`
      - Cleans the variable names
      - Filters the data to include only `pups` with `sex == 1`
      - Removes the `pd_ears` variable
      - Creates a variable that indicates whether `pd_pivot` is 7 or more days
```{r}
# Your answer starts here:  
# use pipe operator!
pups_piped <- read_csv("data/FAS_pups.csv") %>%
  janitor::clean_names() %>%
  filter(sex == 1) %>%
  select(-pd_ears) %>%
  mutate(
    pd_pivot_greater_7 = pd_pivot >= 7
) 
pups_piped
```
  
```{r}
# practice on select
starwars %>% head(6)
starwars %>% select(name:mass)

# more about filter

# keep only the rows where at least one column has a non-missing value
starwars %>% 
  filter(if_any(everything(), ~ !is.na(.)))

# keep only the rows where all columns have non-missing values
starwars %>% 
  filter(if_all(everything(), ~ !is.na(.)))

```
  














